### 2020-08-05 article/highlights.tex

\documentclass[review,authoryear,3p]{elsarticle}
\journal{Remote Sensing of Environment}
\begin{document}
\section*{Highlights}
\begin{itemize}
    \item 3 machine learning model groups tested for deriving 7 land cover fractions globally
    \item Random forest regression performed the best
    \item Optical remote sensing data was the most important input for model predictions
    \item Independent validation showed highest accuracies for built-up and water fractions
    \item Multi-fraction land cover mapping shown as an alternative to discrete classification
\end{itemize}

\end{document}