### 2020-11-18 article/highlights.tex

\documentclass[review,authoryear,3p]{elsarticle}
\journal{Remote Sensing of Environment}
\begin{document}
\section*{Highlights}
\begin{itemize}
    \item 4 machine learning model groups tested for deriving 7 land cover fractions globally
    \item Multi-step approach proposed for dealing with zero inflation in fraction data
    \item Random forest regression performed the best
    \item Optical remote sensing data was the most important input for model predictions
    \item Multi-fraction land cover mapping shown as an alternative to discrete classification
\end{itemize}

\end{document}