### 2021-02-04 article/highlights.tex
Move parts to appendix

\documentclass[review,authoryear,3p]{elsarticle}
\journal{Remote Sensing of Environment}
\begin{document}
\section*{Highlights}
\begin{itemize}
    %\item 4 machine learning model groups tested for deriving 7 land cover fractions globally
    \item 7 land cover fractions derived globally using machine learning algorithms
    \item Multi-step approach proposed for dealing with zero inflation in fraction data
    %\item 1234567890123456789012345678901234567890123456789012345678901234567890123456789012345
    \item The approach provides a middle ground between discrete and fractional mapping
    \item The approach gives the best guess and improves estimation of 0\% and 100\% fractions
    \item Comprehensive spatial accuracy assessment based on over 20\,000 points
    \item Multi-fraction land cover mapping shown as an alternative to discrete classification
\end{itemize}

\end{document}