### 2021-02-18 article/highlights.tex
Final updates for second revision round

\documentclass[review,authoryear,3p]{elsarticle}
\journal{Remote Sensing of Environment}
\begin{document}
\section*{Highlights}
\begin{itemize}
    \item 7 land cover fractions derived using machine learning models and validated globally
    \item Multi-step approach proposed for dealing with zero inflation in fraction data
    \item Independent spatial accuracy assessment based on over 20\,000 reference points
   %\item 1234567890123456789012345678901234567890123456789012345678901234567890123456789012345
    \item Multi-step approach gives a best guess and improves 0\% and 100\% fraction estimation
    \item Multi-fraction land cover mapping shown as an alternative to discrete classification
\end{itemize}

\end{document}